Doubly-Robust Inference in R using \\texttt{drtmle}

### 2022-06-25 paper/using_drtmle_obstudies.Rmd
complete

Suppose the observed data consist of $n$ independent and identically distributed
copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is a vector of
baseline covariates, $A \in \{0,1\}$ is a binary treatment, $Y$ is an outcome,
and $P_0$ is the true data-generating distribution. The parameters of interest
are
\begin{equation}
  \psi_0(a) = E_0\{\bar{Q}(a,W)\} = \int \bar{Q}_0(a,w) dQ_{W,0}(w) \ , \
  \mbox{for} \ a = 0,1 \ ,
  \label{parameterDef}
\end{equation}
where $\bar{Q}_0(a,w)=E_0 \left(Y \mid A=a, W=w\right)$ is the so-called
outcome regression and $Q_{W,0}(w)=Pr_0\left(W\leq w\right)$ is the distribution
function of the covariates. The value $\psi_0(a)$ represents the marginal
(i.e., population-level) treatment-specific average outcome, hereafter referred
to as the *marginal mean* under treatment $A=a$. The marginal effect of the
treatment on the outcome can be assessed by comparing marginal means under
different treatment levels. For example, the average treatment effect is often
defined as $\psi_0 \coloneqq \psi_0(1) - \psi_0(0)$. Under additional
assumptions, these parameters have interpretations as mean counterfactual
outcomes under the different treatments, with contrasts between these means
defining causal effects.

@@

An estimator of $\psi_0(a)$ may be motivated directly by (\ref{parameterDef}).
Suppose we have available an estimate of the OR, $\bar{Q}_n$, and an estimate of
the distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into (\ref{parameterDef}),
$\psi_n(a) = \int \bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical
distribution function of covariates is used so that this estimator can be
equivalently expressed
\begin{equation}
  \psi_{n,\text{GCOMP}}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i)
  \ . \label{eqn:gcomp}
\end{equation}
This estimator is commonly referred to as the G-computation (GCOMP) estimator.
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator (e.g., undersmoothing as discussed above).

@@

An alternative estimator of the treatment-specific population-level mean may be
obtained by noting that (\ref{eqn:gcomp}) can be equivalently written as
\begin{align}
  E_0\{\bar{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
  \label{iptwID}
\end{align}
We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
score. Suppose we had available $g_n$, an estimate of the PS. Equation
(\ref{iptwID}) motivates the estimator
\begin{equation*}
  \psi_{n,\text{IPTW}}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
    \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
\end{equation*}
which is referred to as the inverse probability of treatment weighted (IPTW)
estimator. Like GCOMP estimators, inference for IPTW estimators based on
parametric PS estimators may be facilitated through the nonparametric bootstrap;
however, unlike their GCOMP counterparts, IPTW estimators, being Z-estimators,
are constructed as solutions to an IPTW estimating equation, the empirical
variance of which may also be used for standard error estimation. If adaptive
approaches are used to estimate the PS, inference may not be available without
further modification of the propensity score estimator, whether via targeting,
as we discuss below, or undersmoothing, as explored by
@ertefaie2022nonparametric and @hejazi2022efficient.

@@

The GCOMP and IPTW estimators suffer from two important shortcomings. The first
is a lack of robustness. That is, validity of GCOMP estimators relies on
consistent estimation of the OR, while validity of the IPTW relies on
consistent estimation of the PS. In practice, we may not know which of the OR
or PS is simpler to consistently estimate and thus may not know which of the
two estimators to prefer. Furthermore, if adaptive estimators (e.g., based on
machine learning) are used to estimate the OR or PS, then the GCOMP and IPTW
estimators generally lack a valid basis for asymptotic statistical inference.

This motivates a new class of estimators that combine both the GCOMP and IPTW
estimators. One example is the augmented IPTW (AIPTW) estimator
\begin{equation*}
\psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{IPTW}}(a) + \frac{1}{n} \sum_{i=1}^n
  \biggl\{1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\} \ \bar{Q}_n(a, W_i)
  \ ,
\end{equation*}
which adds an augmentation term to the IPTW estimator. Equivalently, the
estimator can be viewed as adding an augmentation to the GCOMP estimator
\begin{equation} \label{oneStepEst}
\psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{GCOMP}}(a) + \frac{1}{n} \sum_{i=1}^n
\frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i - \bar{Q}_n(a,W_i)\} \ .
\end{equation}
The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
estimators. The estimator exhibits _double robustness_ in that it is consistent
for $\psi_0(a)$ if either $\bar{Q}_n$ is consistent for the true OR *or* $g_n$
is consistent for the true PS. Furthermore, if both the OR and PS are
consistently estimated and regularity conditions hold, one can establish a
normal limiting distribution for the AIPTW estimator. Simple estimators of the
variance of this limiting distribution can be derived that may be used to
construct Wald-style confidence intervals and hypothesis tests.

A more recent proposal to improve robustness and inferential properties of the
GCOMP and IPTW estimators utilizes targeted minimum loss estimation (or TMLE).
TMLE is a general methodology that may be used to modify an estimator of a
regression function in a manner that ensures that the modified estimator
satisfies user-selected equations. In the present problem, one can show that if
an estimator of the OR $\bar{Q}_n^{\star}$ satisfies
\begin{equation}\label{meanEIFZero}
  \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \bar{Q}_n^{\star}(a, W_i)\} = 0 \ ,
\end{equation}
then the GCOMP estimator based on $\bar{Q}_n^{\star}$ will have the same
asymptotic properties as the AIPTW estimator. That is, the estimator will be
doubly robust and, provided both the OR and PS are consistently estimated and
regularity conditions are satisfied, will have a normal limiting distribution.
TMLE provides a means of mapping an initial estimator of the OR into an
estimator that satisfies (\ref{oneStepEst}). In addition to possessing the
desirable asymptotic properties, TML estimators have been shown to outperform
AIPTW estimators in finite samples [@porter:2011:ijb]. We refer readers to
@tlbook:2011 and @vdl2022targeted for further details about on the TMLE
methodology.

@@

@vdl:2014:ijb demonstrated that the double robustness property of AIPTW
and TML estimators does not extend to their normal limiting distribution if
adaptive estimators of the OR and PS are used to construct the estimators.
However, in settings with high-dimensional and continuous-valued covariates,
adaptive estimation is likely necessary in order to obtain a consistent
estimate of either the OR or PS, and thus may be necessary to obtain minimally
biased estimates of the marginal means of interest. @vdl:2014:ijb derived
modified TML estimators that are doubly robust not only with respect to
consistency, but also with respect to asymptotic normality. Furthermore, this
work provided closed-form, doubly robust variance estimators that may be used to
construct doubly robust confidence intervals and hypothesis tests.
@benkeser:2017:biometrika proposed simplifications of the @vdl:2014:ijb
estimators and demonstrated the practical performance of these estimators via
simulation experiments.

The key to the theory underlying these results is finding estimators of the OR
and PS that simultaneously satisfy equation (\ref{oneStepEst}) in addition to
two additional equations. These equations involve additional regression
parameters that we refer to as the reduced outcome regression (R-OR) and the
reduced propensity score (R-PS), defined respectively as
\begin{align}
  \bar{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \bar{Q}_n(W) \mid A = a,
  g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \label{eqn:qbarr} \\
  g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \bar{Q}_n(W)=\bar{Q}_n(w),
  g_n(W)=g_n(w)\right\} \ . \label{eqn:gr_bivariate}
\end{align}
In words, the R-OR is the regression of the residual from the outcome
regression onto the estimated PS amongst observations with $A = a$, while the
R-PS is the propensity for treatment $A = a$ given the estimated OR and PS. We
note that a key feature of these reduced-dimension regressions is that they are
low-dimensional regardless of the dimension of $W$ and depend on neither the
true OR nor the true PS. Thus, these regressions may be consistently estimated
at reasonably fast rates irrespective of the dimension of $W$ and irrespective
of whether the OR or PS are consistently estimated. Based on these
reduced-dimension regressions, @{vdl:2014:ijb showed that if estimates of the
OR $\bar{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR $\bar{Q}_{r,n}^{\star}$, and
R-PS $g_{r,n}^{\star}$ satisfied
\begin{align}
& \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n^{\star}(a \mid W_i)}
\{Y_i - \bar{Q}_n^{\star}(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
&\frac{1}{n} \sum\limits_{i=1}^n \frac{\bar{Q}_{r,n}(a, W_i)}{g_n(a \mid W_i)}
\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
\label{tmleAsolve} \\ &\frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_{r,n}^{\star}(a \mid W_i)}
\biggl\{\frac{g_{r,n}^{\star}(a \mid W_i) - g_n^{\star}(a \mid W_i)}
{g_n^{\star}(a \mid W_i)} \biggr\} \{Y_i - \bar{Q}_n^{\star}(a, W_i) \} = 0 \ ,
\label{tmleQsolve}
\end{align}
then the GCOMP estimator based on $\bar{Q}_n^{\star}$ would be doubly robust
not only with respect to consistency, but also with respect to its limiting
distribution. The author additionally proposed a TMLE algorithm to map initial
estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy these key
equations.

@benkeser:2017:biometrika demonstrated alternative equations that can be
satisfied to achieve this form of double-robustness. In particular, these
authors formulated an alternative version of the R-PS,
\begin{align*}
  g_{r,0,1}(a \mid w) \coloneqq Pr_0\{A = a \mid \bar{Q}_n(W) =
  \bar{Q}_n(w) \} \ ,
\end{align*}
which we refer to as R-PS1. They also introduced an additional regression of a
weighted residual from the PS on the OR,
\begin{align*}
  g_{r,0,2}(a \mid w) \coloneqq E_0\biggl\{\frac{I(A = a) - g_n(a \mid W)}
  {g_n(a \mid W)} \ \bigg\vert \ \bar{Q}_n(W) = \bar{Q}_n(w) \biggr\} \ ,
\end{align*}
which we refer to as R-PS2. The key feature of R-PS1 and R-PS2 is that they are
both univariate regressions, while R-PS is a bivariate regression.
@benkeser:2017:biometrika suggested that this may make R-PS1 and R-PS2
easier to estimate consistently than R-PS. Further, they showed that if
estimators of the OR, PS, and R-OR satisfy equations (5)-(6), and if,
additionally, estimators of the OR, R-PS-1 $g_{r,n,1}^{\star}$, and R-PS-2
$g_{r,n,2}^{\star}$ satisfy
\begin{equation*}
\frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_{r,n,1}^{\star}
(a \mid W_i)} g_{r,n,2}^{\star}(a \mid W_i) \ \{Y_i -
\bar{Q}_n^{\star}(a,W_i)\} = 0 \ ,
\end{equation*}
then the GCOMP estimator based on this $\bar{Q}_n^{\star}$ also enjoys a
doubly-robust limiting distribution. @benkeser:2017:biometrika provided a
TMLE algorithm that produced such estimates and provided closed-form,
doubly-robust variance estimates.

\textbf{Remark: } In the previous section, we introduced AIPTW and TMLE as two
methods for constructing doubly-robust (with respect to consistency) estimators
of our parameter of interest. Theoretically, these two frameworks are closely
related and, practically, the two estimators are generally seen as competitors
for estimating marginal means. In equation (3), we were able to generate a
doubly-robust estimator by adding a term to the GCOMP estimator. This term is
exactly the left-hand-side of the key equation (4), the key equation that is
solved by the standard TMLE. This is no coincidence; this term is crucial in
studying the asymptotic properties of the GCOMP, AIPTW, and TML estimators
(for discussion, see @benkeser:2017:biometrika Section 2.2). Recall that
a TML estimator achieves doubly-robust consistency by satisfying equation (4),
while the AIPTW achieves doubly-robust consistency by combining the GCOMP
estimator with the left-hand-side of this term. Now, we have argued that a TML
estimator may achieve doubly-robust limiting behavior if it additionally
satisfies equations (6) and (7). It is thus sensible to wonder whether an
AIPTW-style estimator could be constructed by combining the left-hand-side of
these two additional equations with the AIPTW estimator.
@benkeser:2017:biometrika showed that, surprisingly, this AIPTW-style estimator
does not achieve the same doubly-robust properties as the TML estimator. Thus,
while AIPTW and TMLE are generally competitors for producing estimators
doubly-robust with respect to consistency, for doubly-robust inference, TMLE
ought to be preferred.

# Implementation of doubly-robust inference

The main function of the package is the eponymous `drtmle()` function. This
function estimates the treatment-specific marginal mean for user-specified
levels of a discrete-valued treatment and computes a doubly-robust covariance
matrix for these estimates. The function implements either the
estimators of @vdl:2014:ijb (if `reduction = "bivariate"`) or the
improved estimators of @benkeser:2017:biometrika
(`reduction = "univariate"`). The package includes utility functions for
combining these estimates into estimates of average treatment effects, as well
as other contrasts, as discussed below.

The main data arguments for `drtmle()` are `W, A, Y`, which, as above, represent
the covariates, a discrete-valued treatment, and an outcome, respectively.
Missing data are allowed in `A` and `Y`, as discussed below. Beyond the data
arguments, it is necessary to specify how to estimate each of the OR, PS, R-OR,
and R-PS (or R-PS1, R-PS2). The package provides flexibility in how these
regression parameters may be estimated. Once the user has specified the data
arguments and how to estimate each regression parameter, the function proceeds
as follows:
  1. estimate the OR via the internal `estimateQ()` function;
  2. estimate the PS via the internal `estimateG()` function;
  3. estimate reduced-dimension regression via the internal `estimateQrn()` and
     `estimategrn()` functions;
  4. iteratively modify initial estimates of the OR and PS via the
     `fluctuateQ()` and `fluctuateG()` functions until equations (5)-(7) are
     approximately solved;
  5. compute the TML estimates and covariance estimates based on the final OR.
We refer interested readers to @benkeser:2017:biometrika for a
theoretical derivation of how the iterative modification of initial OR and PS is
performed.

In addition to returning doubly-robust parameter and covariance estimates, the
`drtmle()` function returns estimates of the GCOMP, AIPTW, standard TMLE, and the
modified AIPTW estimator discussed in the remark above. While doubly-robust
inference based on these estimators is not available, the estimators are
provided for comparison. In the following subsections, we look at various
options for making calls to `drtmle()`.

## Estimating regressions
```{r, echo = FALSE, message = FALSE}
library(drtmle)
set.seed(1234)
N <- 1e6
W <- data.frame(W1 = runif(N), W2 = rbinom(N, 1, 0.5))
Y1 <- rbinom(N, 1, plogis(W$W1 + W$W2))
Y0 <- rbinom(N, 1, plogis(W$W1))
EY1 <- mean(Y1)
EY0 <- mean(Y0)
```
While the estimators computed by `drtmle()` are consistent and asymptotically
normal under inconsistent estimation of either the OR or PS, it is nevertheless
advisable to strive for consistent estimation of *both* the OR and PS. If both
of the regression parameters are consistently estimated, then the estimators
computed by `drtmle()` achieve the nonparametric efficiency bound, resulting in
narrower confidence intervals and more powerful hypothesis tests. In order for
the estimates of the OR and PS to have the greatest chance of consistency, the
`drtmle()` function facilitates estimation using adaptive estimation techniques.
In fact, there are three ways to estimate each of the regression parameters:
generalized linear models, super learning [@vdl:2007:statapp;
@SuperLearnerPackage], and a user-specified estimation technique. We demonstrate
each of these approaches in turn on a  simulated data set. The true parameter
values of interest are $\psi_0(0)=$ `r round(EY0, 2)` and $\psi_0(1)=$ `r
round(EY1, 2)`.

@@

## Additional methods included in drtmle

@@

# Acknowledgements

NSH is grateful for the support of the National Science Foundation (award no.
DMS 2102840).

>>>>

Suppose the observed data consist of $n$ independent and identically distributed
copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is a vector of
baseline covariates, $A \in \{0,1\}$ is a binary treatment, $Y$ is an outcome,
and $P_0$ is the true data-generating distribution. The parameters of interest
are
\begin{equation}
  \psi_0(a) = E_0\{\overline{Q}(a,W)\} = \int \overline{Q}_0(a,w) dQ_{W,0}(w)
  \ , \ \mbox{for} \ a = 0,1 \ ,
  \label{parameterDef}
\end{equation}
where $\overline{Q}_0(a,w)=E_0 \left(Y \mid A=a, W=w\right)$ is the so-called
outcome regression and $Q_{W,0}(w)=Pr_0\left(W\leq w\right)$ is the distribution
function of the covariates. The value $\psi_0(a)$ represents the marginal
(i.e., population-level) treatment-specific average outcome, hereafter referred
to as the *marginal mean* under treatment $A=a$. The marginal effect of the
treatment on the outcome can be assessed by comparing marginal means under
different treatment levels. For example, the average treatment effect is often
defined as $\psi_0 \coloneqq \psi_0(1) - \psi_0(0)$. Under additional
assumptions, these parameters have interpretations as mean counterfactual
outcomes under the different treatments, with contrasts between these means
defining causal effects.

@@

An estimator of $\psi_0(a)$ may be motivated directly by (\ref{parameterDef}).
Suppose we have available an estimate of the OR, $\overline{Q}_n$, and an estimate of
the distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into (\ref{parameterDef}),
$\psi_n(a) = \int \overline{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical
distribution function of covariates is used so that this estimator can be
equivalently expressed
\begin{equation}
  \psi_{n,\text{GCOMP}}(a) = \frac{1}{n}\sum\limits_{i=1}^n
  \overline{Q}_n(a, W_i) \ .
  \label{eqn:gcomp}
\end{equation}
This estimator is commonly referred to as the G-computation (GCOMP) estimator.
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator (e.g., undersmoothing as discussed above).

@@

An alternative estimator of the treatment-specific population-level mean may be
obtained by noting that (\ref{eqn:gcomp}) can be equivalently written as
\begin{align}
  E_0\{\overline{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
  \label{iptwID}
\end{align}
We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
score. Suppose we had available $g_n$, an estimate of the PS. Equation
(\ref{iptwID}) motivates the estimator
\begin{equation*}
  \psi_{n,\text{IPTW}}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
    \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
\end{equation*}
which is referred to as the inverse probability of treatment weighted (IPTW)
estimator. Like GCOMP estimators, inference for IPTW estimators based on
parametric PS estimators may be facilitated through the nonparametric bootstrap;
however, unlike their GCOMP counterparts, IPTW estimators, being Z-estimators,
are constructed as solutions to an IPTW estimating equation, the empirical
variance of which may also be used for standard error estimation. If adaptive
approaches are used to estimate the PS, inference may not be available without
further modification of the propensity score estimator, whether via targeting,
as we discuss below, or undersmoothing, as explored by
@ertefaie2022nonparametric and @hejazi2022efficient.

@@

The GCOMP and IPTW estimators suffer from two important shortcomings. The first
is a lack of robustness -- that is, validity of GCOMP estimators relies on
consistent estimation of the OR, while validity of the IPTW relies on
consistent estimation of the PS. In practice, we may not know which of the OR
or PS is simpler to consistently estimate and thus may not know which of the
two estimators to prefer. Furthermore, if adaptive estimators (e.g., based on
machine learning) are used to estimate the OR or PS, then the GCOMP and IPTW
estimators generally lack a valid basis for asymptotic statistical inference.

This motivates a new class of estimators that combine both the GCOMP and IPTW
estimators. One prominent example is the augmented IPTW (AIPTW) estimator
\begin{equation*}
  \psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{IPTW}}(a) + \frac{1}{n}
  \sum\limits_{i=1}^n \biggl\{1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\}
  \overline{Q}_n(a, W_i) \ ,
\end{equation*}
which adds an augmentation term to the IPTW estimator. Equivalently, the
estimator can be viewed as adding an augmentation to the GCOMP estimator
\begin{equation} \label{oneStepEst}
  \psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{GCOMP}}(a) + \frac{1}{n}
  \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \overline{Q}_n(a,W_i)\} \ .
\end{equation}
The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
estimators: It exhibits _double robustness_ in that it is consistent for
$\psi_0(a)$ if either $\overline{Q}_n$ is consistent for the true OR *or* $g_n$ is
consistent for the true PS. Furthermore, if both the OR and PS are consistently
estimated and regularity conditions hold, one can establish a normal limiting
distribution for the AIPTW estimator. Closed-form estimators of the variance of
this limiting distribution can be derived that may be used to construct
Wald-style confidence intervals and hypothesis tests.

A more recent proposal to improve robustness and inferential properties of the
GCOMP and IPTW estimators utilizes targeted minimum loss estimation (or TMLE).
TMLE is a general methodology that may be used to modify an estimator of a
regression function in a manner that ensures that the modified estimator
satisfies user-selected equations. In the present problem, one can show that if
an estimator of the OR $\overline{Q}_n^{\star}$ satisfies
\begin{equation}\label{meanEIFZero}
  \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \overline{Q}_n^{\star}(a, W_i)\} = 0 \ ,
\end{equation}
then the GCOMP estimator based on $\overline{Q}_n^{\star}$ will have the same
asymptotic properties as the AIPTW estimator. That is, the estimator will be
doubly robust and, provided both the OR and PS are consistently estimated and
regularity conditions are satisfied, will have a normal limiting distribution.
TMLE provides a means of mapping an initial estimator of the OR into an
estimator that satisfies (\ref{oneStepEst}). In addition to possessing the
desirable asymptotic properties, TML estimators have been shown to outperform
AIPTW estimators in finite samples [@porter:2011:ijb]. We refer readers to
@tlbook:2011 and @vdl2022targeted for further details about on the TMLE
methodology.

@@

@vdl:2014:ijb demonstrated that the double robustness property of AIPTW
and TML estimators does not extend to their normal limiting distribution if
adaptive estimators of the OR and PS are used to construct the estimators.
However, in settings with high-dimensional and continuous-valued covariates,
adaptive estimation is likely necessary in order to obtain a consistent
estimate of either the OR or PS, and thus may be necessary to obtain minimally
biased estimates of the marginal means of interest. @vdl:2014:ijb derived
modified TML estimators that are doubly robust not only with respect to
consistency, but also with respect to asymptotic normality. Furthermore, this
work provided closed-form, doubly robust variance estimators that may be used to
construct doubly robust confidence intervals and hypothesis tests.
@benkeser:2017:biometrika proposed simplifications of @vdl:2014:ijb's
estimators and demonstrated the practical performance of these estimators via
simulation experiments.

The key to the theory underlying these results is finding estimators of the OR
and PS that simultaneously satisfy equation (\ref{oneStepEst}) in addition to
two additional equations. These equations involve additional regression
parameters that we refer to as the reduced outcome regression (R-OR) and the
reduced propensity score (R-PS), defined respectively as
\begin{align}
  \overline{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \overline{Q}_n(W) \mid
  A = a, g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \label{eqn:qoverliner} \\
  g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \overline{Q}_n(W) =
  \overline{Q}_n(w), g_n(W)=g_n(w)\right\} \ .
  \label{eqn:gr_bivariate}
\end{align}
In words, the R-OR is the regression of the residual from the outcome
regression onto the estimated PS amongst observations with $A = a$, while the
R-PS is the propensity for treatment $A = a$ given the estimated OR and PS. We
note that a key feature of these reduced-dimension regressions is that they are
low-dimensional regardless of the dimension of $W$ and depend on neither the
true OR nor the true PS. Thus, these regressions may be consistently estimated
at reasonably fast rates irrespective of the dimension of $W$ and irrespective
of whether the OR or PS are consistently estimated. Based on these
reduced-dimension regressions, @vdl:2014:ijb showed that if estimates of the
OR $\overline{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR
$\overline{Q}_{r,n}^{\star}$, and R-PS $g_{r,n}^{\star}$ satisfied
\begin{align}
& \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n^{\star}(a \mid W_i)}
\{Y_i - \overline{Q}_n^{\star}(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
&\frac{1}{n} \sum\limits_{i=1}^n \frac{\overline{Q}_{r,n}(a, W_i)}
{g_n(a \mid W_i)}\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
\label{tmleAsolve} \\ &\frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_{r,n}^{\star}(a \mid W_i)}
\biggl\{\frac{g_{r,n}^{\star}(a \mid W_i) - g_n^{\star}(a \mid W_i)}
{g_n^{\star}(a \mid W_i)} \biggr\} \{Y_i - \overline{Q}_n^{\star}(a, W_i) \}
= 0 \ ,
\label{tmleQsolve}
\end{align}
then the GCOMP estimator based on $\overline{Q}_n^{\star}$ would be doubly
robust not only with respect to consistency, but also with respect to its
limiting distribution. The author additionally proposed a TMLE algorithm to map
initial estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy
these key equations.

@benkeser:2017:biometrika demonstrated alternative equations that can be
satisfied to achieve this form of double-robustness. In particular, these
authors formulated an alternative version of the R-PS,
\begin{align*}
  g_{r,0,1}(a \mid w) \coloneqq Pr_0\{A = a \mid \overline{Q}_n(W) =
  \overline{Q}_n(w) \} \ ,
\end{align*}
which we refer to as R-PS1. They also introduced an additional regression of a
weighted residual from the PS on the OR,
\begin{align*}
  g_{r,0,2}(a \mid w) \coloneqq E_0\biggl\{\frac{I(A = a) - g_n(a \mid W)}
  {g_n(a \mid W)} \ \bigg\vert \ \overline{Q}_n(W) = \overline{Q}_n(w) \biggr\} \ ,
\end{align*}
which we refer to as R-PS2. The key feature of R-PS1 and R-PS2 is that they are
both univariate regressions, while R-PS is a bivariate regression.
@benkeser:2017:biometrika suggested that this may make R-PS1 and R-PS2
easier to estimate consistently than R-PS. Further, they showed that if
estimators of the OR, PS, and R-OR satisfy equations
(\ref{meanEIFZero})--(\ref{eqn:qoverliner}), and if, additionally, estimators of
the OR, R-PS-1 $g_{r,n,1}^{\star}$, and R-PS-2 $g_{r,n,2}^{\star}$ satisfy
\begin{equation*}
\frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_{r,n,1}^{\star}
(a \mid W_i)} g_{r,n,2}^{\star}(a \mid W_i) \ \{Y_i -
\overline{Q}_n^{\star}(a,W_i)\} = 0 \ ,
\end{equation*}
then the GCOMP estimator based on this $\overline{Q}_n^{\star}$ also enjoys a
doubly robust limiting distribution. @benkeser:2017:biometrika provided a
TMLE algorithm that produced such estimates and provided closed-form,
doubly robust variance estimates.

\textbf{Remark: } In the previous section, we introduced AIPTW and TMLE as two
methods for constructing doubly robust (with respect to consistency) estimators
of our parameter of interest. Theoretically, these two frameworks are closely
related and, practically, the two estimators are generally seen as competitors
for estimating marginal means. In equation (\ref{oneStepEst}), we were able to
generate a doubly robust estimator by adding a term to the GCOMP estimator. This
term is exactly the left-hand side of the key equation (\ref{meanEIFZero}), the
key equation
that is solved by the standard TMLE. This is no coincidence -- this term is
crucial in studying the asymptotic properties of the GCOMP, AIPTW, and TML
estimators (for discussion, see @benkeser:2017:biometrika, Section 2.2). Recall
that a TML estimator achieves doubly-robust consistency by satisfying equation
(\ref{meanEIFZero}), while the AIPTW achieves doubly robust consistency by
combining the GCOMP estimator with the left-hand side of this term. Now, we have
argued that a TML estimator may achieve doubly robust limiting behavior if it
additionally satisfies equations (\ref{eqn:qoverliner}) and
(\ref{eqn:gr_bivariate}). It is thus sensible to wonder whether an AIPTW-style
estimator could be constructed by combining the left-hand side of these two
additional equations with the AIPTW estimator. @benkeser:2017:biometrika showed
that, surprisingly, this AIPTW-style estimator does not achieve the same
doubly robust properties as the TML estimator. Thus, while AIPTW and TMLE are
generally competitors for producing estimators doubly robust with respect to
consistency, for doubly robust inference, TML estimators are preferable.

# Implementation of doubly robust inference

The main workhorse of the \texttt{drtmle} package is the eponymous `drtmle()`
function. This function estimates the treatment-specific marginal mean for
user-specified levels of a discrete-valued treatment and computes a doubly
robust covariance matrix for these estimates. The function implements either the
estimators of @vdl:2014:ijb (if `reduction = "bivariate"`) or the improved
estimators of @benkeser:2017:biometrika (if `reduction = "univariate"`). The
package includes utility functions for combining these estimates into estimates
of average treatment effects, as well as other contrasts, as discussed below.

The main data arguments for `drtmle()` are `W, A, Y`, which, as above, represent
the covariates, a discrete-valued treatment, and an outcome, respectively.
Missing data are allowed in `A` and `Y`, as discussed below. Beyond the data
arguments, it is necessary to specify how to estimate each of the OR, PS, R-OR,
and R-PS (or R-PS1 and R-PS2). The package provides flexibility in how these
regression parameters may be estimated. Once the user has specified the data
arguments and how to estimate each regression parameter, the function proceeds
as follows:
  1. estimate the OR via the internal `estimateQ()` function;
  2. estimate the PS via the internal `estimateG()` function;
  3. estimate reduced-dimension regression via the internal `estimateQrn()` and
     `estimategrn()` functions;
  4. iteratively modify initial estimates of the OR and PS via the
     `fluctuateQ()` and `fluctuateG()` functions until equations
     (\ref{meanEIFZero})--(\ref{eqn:gr_bivariate}) are approximately solved;
  5. compute the TML estimates and covariance estimates based on the final OR.
We refer interested readers to @benkeser:2017:biometrika for a theoretical
derivation of how the iterative modification of initial OR and PS is performed.

In addition to returning doubly robust parameter and covariance estimates, the
`drtmle()` function returns estimates of the GCOMP, AIPTW, standard TMLE, and the
modified AIPTW estimator discussed in the remark above. While doubly robust
inference based on these estimators is not available, the estimators are
provided for comparison. In the following subsections, we demonstrate the use
of these methods by examining various options for making calls to `drtmle()`.

## Estimating regressions
```{r, echo = FALSE, message = FALSE}
library(drtmle)
set.seed(1234)
N <- 1e6
W <- data.frame(W1 = runif(N), W2 = rbinom(N, 1, 0.5))
Y1 <- rbinom(N, 1, plogis(W$W1 + W$W2))
Y0 <- rbinom(N, 1, plogis(W$W1))
EY1 <- mean(Y1)
EY0 <- mean(Y0)
```
While the estimators computed by `drtmle()` are consistent and asymptotically
normal under inconsistent estimation of either the OR or PS, it is nevertheless
advisable to strive for consistent estimation of *both* the OR and PS. If both
of the regression parameters are consistently estimated, then the estimators
computed by `drtmle()` achieve the nonparametric efficiency bound (that is,
the asymptotic variance of the efficient influence function), resulting in
narrower confidence intervals and more powerful hypothesis tests. In order for
the estimates of the OR and PS to have the greatest chance of consistency, the
`drtmle()` function facilitates estimation using adaptive estimation techniques.
In fact, there are three ways to estimate each of the regression parameters:
generalized linear models, super learning [@vdl:2007:statapp;
@SuperLearnerPackage], and a user-specified estimation technique. We demonstrate
each of these approaches in turn on a simulated data set in which the treatment
is not randomized. The true parameter values of interest are $\psi_0(0)=$ `r
round(EY0, 2)` and $\psi_0(1)=$ `r round(EY1, 2)`.

@@

## Additional methods included in \texttt{drtmle}

@@

# Acknowledgements {-}
